Heavy rainfall in Paraguay during the 2015-2016 austral summer: causes and subseasonal-to-seasonal predictive skill

### 2017-12-01 reports/writeup/paper.tex
james edits

%Packages
\usepackage{
	siunitx, % units
	cleveref,% cross-references
	physics, % better physics notation
	url, % for URLs
	amsmath, % for math
	array, % for fixed-width tables
	booktabs, % for nice tables
	nth, % for \nth{90}
}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\graphicspath{{../figs/}}

@@

\begin{figure*}
	\noindent\includegraphics[width=6.5in]{StudyArea}
	\caption{
		Topographical map of the study area.
		Colors indicate $\log_{10}$ of elevation, in \si{\meter}, from the Global Land One-Km Base Elevation project available at \texttt{http://iridl.ldeo.columbia.edu/SOURCES/.NOAA/.NGDC/.GLOBE/.topo/}.
		(L): all of South America.
		The domains of the ``Lower Paraguay River Basin'' and the domain used for weather typing are indicated in black and red, respectively.
		(R): The Paraguay River.
		As for (L), the ``Lower Paraguay River Basin'' is indicated.
		Streamflows shown in \cref{fig:streamflow} were taken from the fours stations indicated.
		The Paraguay River and its tributaries, from the Natural Earth database (\texttt{www.naturalearthdata.com}), are also shown.
		Full station names are: Bah\'ia Negra (Bne); Concepci\'on (Conc); Asunci\'on (Asu); Pilar (Pil).
	}
  \label{fig:study-area}
\end{figure*}

@@

\begin{figure*}	\noindent\includegraphics[width=6.5in]{LaggedRain.pdf}
	\caption{
		Circulation anomalies associated with heavy rainfall (\nth{90} percentile exceedance of area-averaged rainfall in the Lower Paraguay River Basin).
		Lagged composites are shown, by column, for $t = $ \SIlist{-2;-1;0;1}{\day} relative to the date of heavy rainfall.
		Top and middle rows show streamfunction anomalies at \SIlist{850}{\hecto\pascal}, respectively.
		Bottom row shows monthly-mean rainfall anomalies, in units of \si{\milli\meter\per\day}.
	}
  \label{fig:lagged-rain}
\end{figure*}

\begin{figure*}
	\noindent\includegraphics[width=6.5in]{PSI_EOF_Loadings.pdf}
	\caption{
    	Leading EOF loadings of daily NDJF \SI{850}{\hecto\pascal} streamfunction $\psi$ over chosen domain.
	}
  \label{fig:eof-loading}
\end{figure*}

\begin{figure*}
	\noindent\includegraphics[width=6.5in,height=9in,keepaspectratio=true]{WTComposite.pdf}
	\caption{
		Composite anomalies associated with each weather type (WT; column).
		Top row shows streamfunction anomalies at \SIlist{850}{\hecto\pascal}.
		Bottom row shows monthly-mean rainfall anomalies, in units of \si{\milli\meter\per\day}. The relative frequency of occurrence of each weather type (in days) is presented on the top of each column.
	}
	\label{fig:wt-composite}
\end{figure*}

\begin{figure*}
	\noindent\includegraphics[width=6.5in]{RainfallWeatherType}
	\caption{
		Time series of area-averaged rainfall in the Lower Paraguay River Basin (\cref{fig:study-area}) for each day of NDJF  2015-16 (lines and points).
        The weather types corresponding to each day are indicated in both color and in an adjacent text label.
		Dashed lines indicate the climatological 50th, \nth{90}, and 99th percentiles of NDJF area-averaged rain over the Lower Paraguay River basin.
	}
  \label{fig:rain-wt}
\end{figure*}

\begin{figure*}
	\noindent\includegraphics[width=6.5in]{NDJF201516Anomaly.pdf}
	\caption{
    	Monthly composite anomalies observed during NDJF 2015-16.
    	Top and middle rows show streamfunction anomalies at \SIlist{200;850}{\hecto\pascal}, respectively.
      Bottom row shows monthly-mean rainfall anomalies, in units of \si{\milli\meter\per\day}.
	}
  \label{fig:anomalies}
\end{figure*}

\begin{figure*}
	\noindent\includegraphics[width=3.5in]{SeasonalForecast.pdf}
	\caption{
	Seasonal model forecast for probability of DJF \nth{90} percentile exceedance.
	Color indicates the forecast probability of exceeding the \nth{90} percentile of climatological rainfall during DJF 2015-16 -- this is presented as the odds ratio as explained in \cref{eq:odds-ratio}.
	A value greater than 1 indicates that the model forecast greater-than-average odds of rainfall exceeding the \nth{90} percentile.
	Grid cells which observed an exceedance of the \nth{90} percentile of DJF rainfall are outlined in black.
	}
  \label{fig:seas-prob-fcst}
\end{figure*}

\begin{figure*}
	\noindent\includegraphics[width=6.5in]{Chiclet.pdf}
	\caption{
    Chiclet diagram \citep{Carbin:2016fx} of ensemble-mean precipitation anomaly forecasts over the Lower Paraguay River Basin (\cref{fig:study-area}) from ECMWF S2S forecast data, as a function of the forecast target date (horizontal axis) and lead time (vertical axis). Time series of CPC daily mean precipitation over the same area is plotted with $y$-axis inverted.
	}
  \label{fig:chiclet}
\end{figure*}

\begin{figure*}
	\noindent\includegraphics[width=6.4in]{ForecastSkill.pdf}
	\caption{
		MOS-adjusted S2S model forecasts and skill scores for the methods indicated in \cref{tab:mos-methods}.
		Top row shows the heavy rainfall forecast for 1-7 December 2015 as the odds ratio of \cref{eq:odds-ratio} over the target domain.
		A value greater than 1 indicates that the model forecast greater-than-average odds of rainfall exceeding the \nth{90} percentile.
        Second row shows the Ignorance score \cref{eq:ignorance}, with zero indicating a perfect forecast.
		Bottom row shows the 2AFC skill score for each grid cell; a value greater than 50 indicates that the model outperforms climatology.
		For all three rows, the grid cells which experienced a \nth{90} percentile exceedance for 1-7 December 2015 are outlined in black.
	}
  \label{fig:subs-prob-fcst}
\end{figure*}

\begin{figure*}
	\noindent\includegraphics[width=6.4in]{WT_MJO_ENSO.pdf}
	\caption{
		Anomalous probability of occurrence of each weather type concurrent with observance of each MJO phase.
		MJO phases are considered here only when the amplitude $\qty(\sqrt{\text{RMM1}^2 + \text{RMM2}^2})$ is greater than 1.
		When amplitude $< 1$, the phase is defined as neutral (0).
		Plots are shown separately for El Ni\~no, La Ni\~na, and neutral ENSO months, defined by the NINO 3.4 index having a value above 1 or below -1.
		Only values which are significant at $p<0.10$, calculated using a bootstrap with 3000 samples, are shown.
	}
  \label{fig:wt-mjo-enso}
\end{figure*}

@@

\begin{figure*}
	\noindent\includegraphics[width=6.4in]{WT4_Monthly_Composite.pdf}
	\caption{
		Composite sea surface temperature (SST) anomalies of the 15 NDJF months with the most occurrences of weather type 4.
		SST data, in \si{\celsius}, come from \citet{Reynolds:2002iy}.
		To match the SST data, only months since NDJF 1981 are considered.
	}
  \label{fig:wt4-months}
\end{figure*}

>>>>

%Packages
\usepackage{
	siunitx, % units
	cleveref,% cross-references
	physics, % better physics notation
	url, % for URLs
	amsmath, % for math
	array, % for fixed-width tables
	booktabs, % for nice tables
	nth, % for \nth{90}
}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\graphicspath{{../../figures/}{./}}

@@

\begin{figure*}
	\noindent\includegraphics[width=6.5in]{study-area.jpg}
	\caption{
		Topographical map of the study area.
		Colors indicate $\log_{10}$ of elevation, in \si{\meter}, from the Global Land One-Km Base Elevation project available at \texttt{http://iridl.ldeo.columbia.edu/SOURCES/.NOAA/.NGDC/.GLOBE/.topo/}.
		(L): all of South America.
		The domains of the ``Lower Paraguay River Basin'' and the domain used for weather typing are indicated in black and red, respectively.
		(R): The Paraguay River.
		As for (L), the ``Lower Paraguay River Basin'' is indicated.
		Streamflows shown in \cref{fig:streamflow} were taken from the fours stations indicated.
		The Paraguay River and its tributaries, from the Natural Earth database (\texttt{www.naturalearthdata.com}), are also shown.
		Full station names are: Bah\'ia Negra (Bne); Concepci\'on (Conc); Asunci\'on (Asu); Pilar (Pil).
	}
  \label{fig:study-area}
\end{figure*}

@@

\begin{figure*}	\noindent\includegraphics[width=6.5in]{lagged-rainfall.pdf}
	\caption{
		Circulation anomalies associated with heavy rainfall (\nth{90} percentile exceedance of area-averaged rainfall in the Lower Paraguay River Basin).
		Lagged composites are shown, by column, for $t = $ \SIlist{-2;-1;0;1}{\day} relative to the date of heavy rainfall.
		Top and middle rows show streamfunction anomalies at \SIlist{850}{\hecto\pascal}, respectively.
		Bottom row shows monthly-mean rainfall anomalies, in units of \si{\milli\meter\per\day}.
	}
  \label{fig:lagged-rain}
\end{figure*}

\begin{figure*}
	\noindent\includegraphics[width=6.5in]{eof-loadings.pdf}
	\caption{
    	Leading EOF loadings of daily NDJF \SI{850}{\hecto\pascal} streamfunction $\psi$ over chosen domain.
	}
  \label{fig:eof-loading}
\end{figure*}

\begin{figure*}
	\noindent\includegraphics[width=6.5in,height=9in,keepaspectratio=true]{weather-type-composite.pdf}
	\caption{
		Composite anomalies associated with each weather type (WT; column).
		Top row shows streamfunction anomalies at \SIlist{850}{\hecto\pascal}.
		Bottom row shows monthly-mean rainfall anomalies, in units of \si{\milli\meter\per\day}. The relative frequency of occurrence of each weather type (in days) is presented on the top of each column.
	}
	\label{fig:wt-composite}
\end{figure*}

\begin{figure*}
	\noindent\includegraphics[width=6.5in]{wt-rain-time-series.pdf}
	\caption{
		Time series of area-averaged rainfall in the Lower Paraguay River Basin (\cref{fig:study-area}) for each day of NDJF  2015-16 (lines and points).
        The weather types corresponding to each day are indicated in both color and in an adjacent text label.
		Dashed lines indicate the climatological 50th, \nth{90}, and 99th percentiles of NDJF area-averaged rain over the Lower Paraguay River basin.
	}
  \label{fig:rain-wt}
\end{figure*}

\begin{figure*}
	\noindent\includegraphics[width=6.5in]{circulation-NDJF-1516-anomaly.pdf}
	\caption{
    	Monthly composite anomalies observed during NDJF 2015-16.
    	Top and middle rows show streamfunction anomalies at \SIlist{200;850}{\hecto\pascal}, respectively.
      Bottom row shows monthly-mean rainfall anomalies, in units of \si{\milli\meter\per\day}.
	}
  \label{fig:anomalies}
\end{figure*}

\begin{figure*}
	\noindent\includegraphics[width=3.5in]{seasonal-forecast.pdf}
	\caption{
	Seasonal model forecast for probability of DJF \nth{90} percentile exceedance.
	Color indicates the forecast probability of exceeding the \nth{90} percentile of climatological rainfall during DJF 2015-16 -- this is presented as the odds ratio as explained in \cref{eq:odds-ratio}.
	A value greater than 1 indicates that the model forecast greater-than-average odds of rainfall exceeding the \nth{90} percentile.
	Grid cells which observed an exceedance of the \nth{90} percentile of DJF rainfall are outlined in black.
	}
  \label{fig:seas-prob-fcst}
\end{figure*}

\begin{figure*}
	\noindent\includegraphics[width=6.5in]{chiclet-s2s-area-averaged.pdf}
	\caption{
    Chiclet diagram \citep{Carbin:2016fx} of ensemble-mean precipitation anomaly forecasts over the Lower Paraguay River Basin (\cref{fig:study-area}) from ECMWF S2S forecast data, as a function of the forecast target date (horizontal axis) and lead time (vertical axis). Time series of CPC daily mean precipitation over the same area is plotted with $y$-axis inverted.
	}
  \label{fig:chiclet}
\end{figure*}

\begin{figure*}
	\noindent\includegraphics[width=6.4in]{s2s-forecast-mos.pdf}
	\caption{
		MOS-adjusted S2S model forecasts and skill scores for the methods indicated in \cref{tab:mos-methods}.
		Top row shows the heavy rainfall forecast for 1-7 December 2015 as the odds ratio of \cref{eq:odds-ratio} over the target domain.
		A value greater than 1 indicates that the model forecast greater-than-average odds of rainfall exceeding the \nth{90} percentile.
        Second row shows the Ignorance score \cref{eq:ignorance}, with zero indicating a perfect forecast.
		Bottom row shows the 2AFC skill score for each grid cell; a value greater than 50 indicates that the model outperforms climatology.
		For all three rows, the grid cells which experienced a \nth{90} percentile exceedance for 1-7 December 2015 are outlined in black.
	}
  \label{fig:subs-prob-fcst}
\end{figure*}

\begin{figure*}
	\noindent\includegraphics[width=6.4in]{wt-mjo-enso.pdf}
	\caption{
		Anomalous probability of occurrence of each weather type concurrent with observance of each MJO phase.
		MJO phases are considered here only when the amplitude $\qty(\sqrt{\text{RMM1}^2 + \text{RMM2}^2})$ is greater than 1.
		When amplitude $< 1$, the phase is defined as neutral (0).
		Plots are shown separately for El Ni\~no, La Ni\~na, and neutral ENSO months, defined by the NINO 3.4 index having a value above 1 or below -1.
		Only values which are significant at $p<0.10$, calculated using a bootstrap with 3000 samples, are shown.
	}
  \label{fig:wt-mjo-enso}
\end{figure*}

@@

\begin{figure*}
	\noindent\includegraphics[width=6.4in]{ssta-composite-wt4.pdf}
	\caption{
		Composite sea surface temperature (SST) anomalies of the 15 NDJF months with the most occurrences of weather type 4.
		SST data, in \si{\celsius}, come from \citet{Reynolds:2002iy}.
		To match the SST data, only months since NDJF 1981 are considered.
	}
  \label{fig:wt4-months}
\end{figure*}
